---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figures/README-", 
  message = FALSE, warning = FALSE
  )
```

# boeCharts

## Overview

`boeCharts` is an R package to help you create charts that are in-line with [the Bank's visual identity standards](https://bankofengland.frontify.com/d/RPk6pMZziBFw/bank-standards), recommended (and designed) for use in conjunction with [ggplot2](https://ggplot2.tidyverse.org/). It contains design patterns for approximating charts found within flagship publications, including:

- [Bank Overground](https://www.bankofengland.co.uk/bank-overground)
- [Monetary Policy Report](https://www.bankofengland.co.uk/monetary-policy-report/2019/november-2019) (FKA Inflation Report)
- [Financial Stability Report](https://www.bankofengland.co.uk/financial-stability-report/2019/july-2019)
- [Statistical Releases](https://www.bankofengland.co.uk/statistics/money-and-credit/2019/april-2019)

N.B. this is still a maturing project. Feedback/issues/contributions appreciated. If you are reporting a bug/error and/or requesting a new feature, please try and use [the boeCharts backlog](https://almplatform/tfs/UnmanagedCollection/Shared%20Analytical%20Code/_backlogs/backlog/boeCharts/Requirements). For questions and other discussion, please use the [Data Community](https://bankexchange/groups/1067/SitePages/Home.aspx).

## Installation

Ensure you are installing R packages via [Artifactory](https://binarycentral/artifactory/webapp/#/home) by **running this once** (within an R session):

```{r, eval=FALSE}
local({
  r <- list("boe-cran-remote-repo" = "https://binarycentral/artifactory/boe-cran-remote-repo/",
            "boe-cran-local-repo" = "https://binarycentral/artifactory/boe-cran-local-repo/",
            "rcran" = "http://cran.rstudio.com/"
  )
  options(repos = r)
})
```

You can then install the latest stable version of `boeCharts` from Artifactory with:

``` {r, eval=FALSE}
install.packages("boeCharts")
```

You can also install the development version of `boeCharts` like so:

```{r, eval=FALSE}
remotes::install_git(
  "https://almplatform/tfs/UnmanagedCollection/Shared%20Analytical%20Code/_git/boeCharts", 
  git = "external"
)
```

N.B. the development version should only be used for experimentation. Use the latest stable version, released via Artifactory, in any production code.

## Usage

### Complementing a ggplot2 worflow

`boeCharts` provides native support for use with [ggplot2](https://ggplot2.tidyverse.org/) (a popular R graphics library). You can use `scale_[colour|fill]_[continuous|discrete]_boe` functions for adopting Bank colour palettes within charts generated by `ggplot2`, and `theme_xxx` functions for applying full chart styles supported by `boeCharts`.

#### A minimal example

First, load packages:

```{r, message=TRUE}
library(boeCharts)
library(ggplot2)
library(palmerpenguins) # for example purposes only
```

Here is an out-of-the-box `ggplot2` chart.

```{r}
# build chart
(chart <- ggplot(penguins, aes(flipper_length_mm, body_mass_g, color = species)) +
  # jittery scatter
  geom_jitter(size = 3)
)
```

Custom themes can be applied to ggplot objects as additional "layers". As mentioned above, all `boeCharts` theme functions are prefixed by `theme_` e.g. `theme_overground()` applies the [Bank Overground](https://www.bankofengland.co.uk/bank-overground) chart theme:

```{r}
(chart <- chart +
  # add Bank Overground theme
  theme_overground()
)
```

Custom ggplot scales help with applying Bank colour palettes. In this case, the points can be coloured with a "vibrant" Bank colour combination variant, via `scale_colour_discrete_boe()`, again as an additional ggplot layer:
 
```{r}
chart <- chart +
  # add a "vibrant" Bank colour combination
  scale_colour_discrete_boe(palette = "vibrant_c")

chart
```
 
Now, switching in the [Monetary Policy Report](https://www.bankofengland.co.uk/monetary-policy-report/2019/november-2019) theme and pre-MPC colours. Additionally, `scale_y_continuous()` is used to move the y-axis across to the right.

```{r}
chart +
  # add MPR theme
  theme_mpr() +
  # add a "vibrant" Bank colour combination
  scale_colour_discrete_boe(palette = "pre_mpc") +
  # move the y-axis to the right
  scale_y_continuous(
    position = "right", sec.axis = dup_axis(labels = NULL)
    )
```

#### An in-depth example

Another `ggplot2` + `boeCharts` creation, this time investigating some more customization options, including automatic axis breaks/limits (using `boe_breaks|limits_date|numeric()`), direct line labels (using `position_voronoi()`) and above-plot y-axis title positioning (using `move_ylab()`).

```{r, example}
# create chart
chart_fang <- ggplot(data = FANG, aes(x = date, y = close, colour = symbol)) +
  # add lines + hide legend
  geom_line(lwd = 0.75, show.legend = FALSE) +
  # add series labels + hide legend
  geom_text(aes(label = symbol), position = position_voronoi(), 
            family = "Calibri", show.legend = FALSE) +
  # add some chart labels
  labs(
    title = "BoE Palette Test", subtitle = "A plot for demonstration purposes",
    caption = caption_boe(source = "Investopedia"),
    y = "Closing price", x = NULL
    ) +
  # use 'highlights' palette
  scale_colour_discrete_boe(palette = "boe_highlights") +
  # add Bank Overground theme
  theme_overground() +
  # apply custom axis settings
  scale_y_continuous(
    expand = c(0, 0), breaks = boe_breaks_numeric(), 
    limits = boe_limits_numeric(), position = "right"
    ) +
  scale_x_date(
    expand = c(0, 0), labels = boe_date_labels(),
    breaks = boe_breaks_date(), limits = boe_limits_date()
    )

# re-position y axis title above plot
move_ylab(chart_fang)
```

#### Markdown theme variants

A Markdown variant of each {boeCharts} theme has been made with a `md_` suffix, like `theme_fsr_md()`. These themes add support for rendering text as markdown, thanks to the [ggtext](https://wilkelab.org/ggtext/index.html) package. Here's an example:

```{r}
fsr_chart_fang <- chart_fang +
  # apply custom axis settings
  scale_x_date(
    labels = boe_date_labels(),
    breaks = boe_breaks_date()
    ) +
  theme_fsr_md(axis_title_size = 9, axis_text_size = 9, caption_size = 9) +
  labs(title = "**Chart A.2** BoE Palette Test",
       subtitle = "A plot for demonstration purposes<sup>(a)</sup>",
       caption = caption_boe(source = "Investopedia", footnote = "A minor data quibble."))

move_ylab(fsr_chart_fang)
```


### Using custom fonts

`boeCharts` helps you safely import and use fonts other than the basic fonts R already uses (with help from the `extrafont` package). These fonts can be imported using an associated `import_` utility function. For example, to import Calibri (used by default in some `boeCharts` themes), you should run the `import_calibri()` function. If you want to use fonts in PDF (or Postscript) output files, set the `boeCharts.loadfonts` option to `TRUE`.

Font import depends on version 1.3.8 of the {rttf2pt1} package being available. The `install_rttf2pt1()` function is included in {boeCharts} to help handle this, and is run automatically as part of the `import_` operations described above.

N.B. the font import step only needs to be run once on your machine. Fonts will be loaded automatically when you load `boeCharts` in future sessions.

## More guides/examples

Talks:

- [Bank-style data viz made simple](http://intranet/Banknav/IML.asp?svr=BOE-DMS&db=Analytical&id=8093829&v=0)

Vignettes:

- [Using boeCharts](http://collaborate/workspaces/RHelpCentre/R%20Markdown/using-boeCharts.html)
- [Bank-standard colours with boeCharts](http://collaborate/workspaces/RHelpCentre/R%20Markdown/Bank-standard-colours-with-boeCharts.html)
- [‘Last-mile’ formatting](http://collaborate/workspaces/RHelpCentre/R%20Markdown/last-mile-formatting.html)

Cookbooks:

- [Introduction to boeCharts - produce “Bank-ready” data visualisation with ggplot2](http://collaborate/workspaces/RHelpCentre/R%20Markdown/boeCharts_intro.html)
- [Visualising Distributions](http://collaborate/workspaces/RHelpCentre/R%20Markdown/ChartsVisualisingDistributions.html)
- [Other charts from SPERI speech](http://collaborate/workspaces/RHelpCentre/R%20Markdown/ChartsOthers.html)

## How to contribute

This is an ongoing project to abstract and consolidate Bank charting guidelines, and any/all contributions are encouraged. Inspecting the source code in this repository will help with understanding how to make a custom palette, or building a chart theme for your business area. The following resources are also helpful (and inspired much of this effort):

- [Modify components of a theme](https://ggplot2.tidyverse.org/reference/theme.html)
- [govstyle](https://github.com/ukgovdatascience/govstyle)
- [hrbrthemes](https://github.com/hrbrmstr/hrbrthemes)

Also, for general consideration when contributing to the package, try to:

- Observe principles in Hadley Wickham's [R Packages book](https://r-pkgs.org/)
- Follow the Git branch (see [here](https://blog.rstudio.com/2017/09/13/rstudio-v1.1-the-little-things/) for how to create a branch from within R Studio) & [pull request](https://docs.microsoft.com/en-us/azure/devops/repos/git/pullrequest?view=azure-devops) model, or [contact Ewen](mailto:ewen.henderson@bankofengland.co.uk)
