---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.path = "figures/", dpi = 105, fig.width = 6, fig.asp = 0.618
)
```

# boeCharts

## Overview

The `boeCharts` R package contains Bank styles for [ggplot2](https://github.com/hadley/ggplot2) charts. It currently houses designs to help you approximate charts found in these places:

- [BoE website](https://www.bankofengland.co.uk/statistics/visual-summaries/effective-interest-rates)
- [Inflation Report](https://www.bankofengland.co.uk/inflation-report/2018/august-2018/global-economic-and-financial-market-developments)
- [DSD Statistical Publications](https://www.bankofengland.co.uk/-/media/boe/files/statistics/money-and-credit/2018/february-2018.pdf?la=en&hash=D5A6531045C648B4169D5FD480723AE4CFBD75F9)

## Install

**boeCharts** is an R package/library best used via R Studio, installed from the [software catalogue](http://sccm-wl-mgt-01/CMApplicationCatalog).

Get the latest stable version of **boeCharts** by running the below code within an R session:

```{r install, eval=FALSE}
# install remotes (required for installation)
if (!require("remotes")) install.packages("remotes")

# install boeCharts from collaborate
install.packages("http://collaborate/workspaces/RHelpCentre/R%20documents/Packages/boeCharts_1.0.0.zip", repos = NULL, type = "binary")

# install boeCharts' dependencies from CRAN
remotes::install_deps(find.package("boeCharts"))

```

## Use

`boeCharts` provides native support for use with [ggplot2](https://ggplot2.tidyverse.org/) (a popular R graphics library). You can use `scale_[colour|fill]_boe` functions for adopting Bank colour palettes within charts generated by `ggplot2`, and `theme_xxx` functions for applying full chart styles supported by `boeCharts`.

```{r, example}
# load packages
library(boeCharts)
library(ggplot2)

# create chart
ggplot(data = FANG, aes(x = date, y = close, colour = symbol)) +
  geom_line(size = 0.7) +
  labs(title = "BoE Palette Test",
       subtitle = "A plot for demonstration purposes",
       y = "Closing\nprice", x = NULL) +
  # apply boe colour palette
  scale_colour_boe() +
  # apply an MCG publication theme
  theme_mcg_pub(axis_text_size = 11) +
  # apply custom axis settings
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 1250, 250), 
                     limits = c(0, 1250), position = "right") +
  scale_x_date(expand = c(0, 0), labels = boe_date_labels(),
               breaks = seq(min(FANG$date), max(FANG$date), by = "3 months"))
  
```

For more detailed guidance, refer to the package [vignettes](https://tfsapp-liv/tfs/UnmanagedCollection/Shared%20Analytical%20Code/_git/boeCharts?path=%2Fvignettes&version=GBmaster&_a=contents) (also included in the `boeCharts` documentation, accessible from within R once the package is installed).

## Contribute

This is an ongoing project to abstract and consolidate Bank charting guidelines, and any/all contributions are encouraged. Inspecting the source code in this repository will help with understanding how to make a custom palette, or building a chart theme for your business area. The following resources are also helpful (and inspired much of this effort):

- [Modify components of a theme](https://ggplot2.tidyverse.org/reference/theme.html)
- [govstyle](https://github.com/ukgovdatascience/govstyle)
- [hrbrthemes](https://github.com/hrbrmstr/hrbrthemes)
- [Building a new theme](https://bookdown.org/rdpeng/RProgDA/building-a-new-theme.html)

Also, for general consideration when contributing to the package, try to:

- Observe principles in Hadley Wickham's [R Packages book](https://r-pkgs.org/)
- Follow the Git branch (see [here](https://blog.rstudio.com/2017/09/13/rstudio-v1.1-the-little-things/) for how to create a branch from within R Studio) & [pull request](https://docs.microsoft.com/en-us/azure/devops/repos/git/pullrequest?view=azure-devops) model, or [contact Ewen](mailto:ewen.henderson@bankofengland.co.uk)

## TO DO

- Implementation of "colour combinations" described in new identity guidelines
- Bank Overground theme
- Release new minor version
