---
title: "Bank-standard colours with boeCharts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bank-standard colours with boeCharts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

## BoE Visual Identity
In early 2019, Communications directorate [updated the Bank's visual identity](http://intranet/BankNews/Pages/New-yearnew-lookUpdatingt-636840277763400290.aspx) including a suite of standard colours for BoE outputs.  For more details on the colours, including accessibility ratings and general guidance, see the [Brand Standards](https://bankofengland.frontify.com/d/RPk6pMZziBFw/bank-standards#/brand-elements/colour-palette).

We hould use these in our charts.  But how to do it without mucking around with RGB codes and Hex values?  **boeCharts to the rescue!**

### The `boeCharts` package
The first time you use it, you'll need to install it.  This is an internaly built package, so rather than the regular `install.packages("boeCharts")` to pull from CRAN, we specify the full path and `repos=NULL`:

```{r install, eval=FALSE}
install.packages("http://collaborate/workspaces/RHelpCentre/R%20Documents/Packages/boeCharts_1.0.1.zip", repos=NULL, type="binary")
```

and then library it:

```{r lib, message=FALSE, warning=FALSE}
library(boeCharts)
library(dplyr)
```

### What are the colours?
There are a number of palettes we can use:  the standard BoE palette, plus Soft, Highlights, Neutral & Rich.  The `boeCharts` package also includes palettes for other business areas.

We can see these as named vectors in the list `boe_palettes`, or as individual vectors `boe`, `boe_highlights`, etc.

```{r palette names}
boe_palettes %>% names()
```

```{r palette values}
boe_palettes$boe
boe_palettes$harmonious_orange
```

Hex codes are all very well, but we want to choose our colours visually.  To help us we can view the palettes using `show_palette_cols()`:

```{r show palettes}
show_palette_cols(boe_palettes$boe)
title("boe")

show_palette_cols(boe_palettes$boe_highlights)
title("boe highlights")

show_palette_cols(boe_palettes$boe_neutral)
title("boe neutral")

show_palette_cols(boe_palettes$boe_rich)
title("boe rich")

show_palette_cols(boe_palettes$boe_soft)
title("boe soft")

```

To see all the colour palettes, try `show_colour_swatches()`.
```{r colour_swatches, echo=F}
## Not run: launches a PDF colour chart
show_colour_swatches()
```

## Applying BoE colours to charts with `scale_xxx_boe`

How can we apply these palettes to our charts?

Let's take a classic chart, using the iris dataset.  Here's the default output:
```{r first chart, message=FALSE, warning=FALSE}
library(ggplot2)

iris %>% 
  ggplot(aes(Petal.Length, Petal.Width, colour = Species)) +
  geom_point()
```

### `scale_fill_discrete_boe()` and `scale_colour_discrete_boe()`
In the chart above, the colour aesthetic is mapped to `Species`.  So we use **`scale_colour_discrete_boe`** to apply some BoE official colours.

```{r boe point}

iris %>% 
  ggplot(aes(Petal.Length, Petal.Width, colour = Species)) +
  geom_point() +
  scale_colour_discrete_boe()
```

Easy as that.

For other chart types which take a `fill` attribute, use the equivalent `scale_fill_discrete_boe()`.

```{r boe bar, fig.height=4}
iris %>% 
  group_by(Species) %>% 
  summarise(mean_petal_length = mean(Petal.Length)) %>% 
  ggplot(aes(Species, mean_petal_length, fill = Species)) +
  geom_col() +
  scale_fill_discrete_boe()
```



## Variations

There are a number of variations we can make to this, including

* Reversing colour order
* Choosing different palettes
* Picking specific colours

We will look at each of these in turn.


### Reversing the palette
By passing `reverse=TRUE`, ggplot will pick colours from the end of the palette first, rather than the beginning.

```{r reverse}
iris %>% 
  ggplot(aes(Petal.Length, Petal.Width, colour = Species)) +
  geom_point() +
  scale_colour_discrete_boe(reverse = TRUE)

```

### Changing palette

We have [5 primary BoE palettes](https://bankofengland.frontify.com/d/RPk6pMZziBFw/bank-standards#/brand-elements/colour-palette/colour-values) and a number of Harmonious and Vibrant colour combinations to choose from, plus some custom palettes:
```{r palette names2}
print(boe_palettes %>% names())
```

Let's change our chart to the BoE Highlights palette:
```{r diff palette}
iris %>% 
  ggplot(aes(Petal.Length, Petal.Width, colour = Species)) +
  geom_point() +
  scale_colour_discrete_boe(boe_highlights) ## boe, boe_highlights, boe_rich, boe_soft, boe_neutral, harmonious_xxx, vibrant_x

```

### Specifying particular colours

We can pick out our favourite colours from the named vectors, and combine them as we like.  This also lets us mix and match colours from different palettes.

```{r indiv cols}

iris %>% 
  group_by(Species) %>%
  summarise(`Mean Petal Length` = mean(Petal.Length)) %>% 
  ggplot() +
  geom_col(aes(Species, `Mean Petal Length`, fill = Species)) +
  scale_fill_manual(values = c(boe_soft$yellow, boe_highlights$orange, boe_rich$rich_orange))

```

## Continuous palettes
```{r continuous}
## diverging palette from boe_diverging_palettes
iris %>% 
  ggplot() +
  geom_point(aes(Petal.Width, Petal.Length, colour = Petal.Width)) +
  scale_colour_continuous_boe(boe_diverging_palettes$red_blue) +
  theme_minimal()

## specify a particular midpoint
iris %>% 
  ggplot() +
  geom_point(aes(Petal.Width, Petal.Length, colour = Petal.Width)) +
  scale_colour_continuous_boe(boe_diverging_palettes$red_blue, midpoint=2) +
  theme_minimal()

```

## Creating custom palettes using BoE colours

### Continuous
We can use `scale_colour_gradient()` for a sequential palette, and `scale_colour_gradient2()` for a diverging palette, providing the start and end colours (and the mid colour, in the case of `_gradient2`).  The interim colours will be interpolated for us.

```{r continuous custom}
# sequential palette
iris %>% 
  ggplot() +
  geom_point(aes(Petal.Width, Petal.Length, colour = Petal.Width)) +
  scale_colour_gradient(low = boe_neutral$light_grey,
                        high = boe$fuchsia) +
  theme_minimal()

```

### Discrete (categorical)
We can use `colorRampPalette()` to create our own palettes:  we provide a vector of colours for the start, end, and optionally interim points.

```{r}
## Add a factor column, binning the petal widths
iris2 <- 
  iris %>% 
  mutate(Petal.Width.Bin = cut(Petal.Width, seq(0, 3, by = 0.5)))
n_bins <- iris2$Petal.Width.Bin %>% levels %>% length


custom_diverging_palette <- colorRampPalette(c(boe$red, boe$light_green, boe$mid_blue))(n_bins)

iris2 %>% 
  ggplot() +
  geom_point(aes(Petal.Width, Petal.Length, colour = Petal.Width.Bin)) +
  scale_colour_manual(values = custom_diverging_palette) +
  theme_minimal()

custom_sequential_palette <- colorRampPalette(c(boe$yellow, boe$dark_green))(n_bins)

iris2 %>% 
  ggplot() +
  geom_point(aes(Petal.Width, Petal.Length, colour = Petal.Width.Bin)) +
  scale_colour_manual(values = custom_sequential_palette) +
  theme_minimal()
```


*Al Firrell, Advanced Analytics, May 2019*

