---
title: "'Last-mile' formatting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{'Last-mile' formatting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height = 4, cache=FALSE, dev = 'svg', fig.dpi = 96
)
```

Once you've applied `boeCharts` theme and palette approximations (detailed in [using boeCharts](using-boeCharts.html)), there may still be fine-tuning necessary to faithfully produce a publication-ready chart. Functions to assist with these 'last-mile' aesthetic tasks are detailed here, some implemented within this package and some general `ggplot2` pointers.


## Axis labels

#### Dates

There is an established preference in some areas for date labels that include the year if it hasn't appeared on the chart previously i.e. only when a 'new' year appears. This style of date labelling isn implemented with `boe_date_labels()`.

```{r}

library(ggplot2)
library(boeCharts)

# create chart
p <- ggplot(data = FANG, aes(x = date, y = close, colour = symbol)) +
  geom_line() +
  labs(title = "Formatting Tests", subtitle = "A plot for demonstration purposes",
       y = "Closing\nprice", x = NULL)

p +
  # set axis date label style
  scale_x_date(date_breaks = "3 months", labels = boe_date_labels())

```

#### Currencies

When plotting numeric values that represent currencies, it's helpful to denote this while retaining the numeric class of your values. Below is an example of `format_pound()`.

```{r}

p +
  # set axis currency label style
  scale_y_continuous(labels = format_pound())

```


## Axis scales

#### End on tick marks

Make sure your axes start and end on tick marks by explicitly setting the `expand`, `limits` and `breaks` settings in `scale...` functions of `ggplot2`.

```{r}
p +
  # apply MCG theme
  theme_mcg_pub() +
  # apply axis scale settings
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 1250, 250), limits = c(0, 1250), position = "right") +
  scale_x_date(expand = c(0, 0), breaks = seq(min(FANG$date), max(FANG$date), by = "3 months"), labels = boe_date_labels())
```

