---
title: "Using boeCharts"
author: "Ewen Henderson, Data & Statistics"
date: "July 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using boeCharts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", cache = FALSE, fig.dpi = 96, fig.width = 7, fig.height = 5)

```

The `boeCharts` package provides tools and helpers to help craft and productionise "Bank-ready" charts, primarily be augmenting [ggplot2](https://ggplot2.tidyverse.org/) graphics. These includes full chart themes for publications, recognised colour palettes, and helpful 'last-mile' utilities (detailed in ['Last-mile' formatting](last-mile-formatting.html)) for common, smaller tasks like formatting date labels.

## Bank colour palettes

Incorporate Bank palettes into your charts, like those found in [Bank standards on corporate identity](https://bankofengland.frontify.com/d/RPk6pMZziBFw/bank-standards#/brand-elements/colour-palette). 

Check out the options available by calling `palette_names`.

```{r}
# load package
library(boeCharts)

palette_names()
```

You can preview colours in a supported palette using `boe_pal`:

```{r}
boe_pal(palette = "boe", n = 4)
```

The easiest way to use a palette in a chart is to use `scale_[colour|fill]_discrete_boe` as part of a standard `ggplot2` workflow (i.e. use in the same way you would use `scale` functions out-of-the-box in `ggplot2`):

```{r boe_pal}
# load package
library(ggplot2)

# make a base plot
chart <- ggplot(data = FANG, aes(x = date, y = close, colour = symbol)) +
  geom_line() +
  labs(title = "BoE palette & theme test",
       subtitle = "A plot for demonstration purposes",
       y = "Closing\nprice", x = NULL) +
  theme_minimal()

# apply boe palette
chart +
  scale_colour_discrete_boe(palette = "boe")

```

#### Business area / publication-ready palettes

Use colour schemes for your own business area or publications. For example, use colours seen in [historic Money & Credit statistical releases](https://www.bankofengland.co.uk/-/media/boe/files/statistics/money-and-credit/2017/october-2017.pdf?la=en&hash=F91456AC176039665F053EEE634CC2DB6D2DA506) with the "mcg" palette.

```{r mcg_pal}

chart +
  # apply mcg palette
  scale_colour_discrete_boe(palette = "mcg")

```

## Chart themes

Go beyond palettes, and control the appearance of all (non-data) chart elements with **themes** to produce Bank-style plots.

### Bank Overground

`theme_overground` approximates the design of charts in [Bank Overground](https://www.bankofengland.co.uk/bank-overground) publications (note the use of a vibrant colour combination in this example, too).

```{r overground_theme}

chart_two <- ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_jitter(size = 3) +
  # Bank Overground theme
  theme_overground() +
  # a "vibrant" Bank colour combination
  scale_colour_discrete_boe(palette = "vibrant_c") +
  labs(title = "BoE palette & theme test",
       subtitle = "A plot for demonstration purposes")

chart_two
```

The helper function `left_align_titles` can be used to fix titles to the left of the chart window (as is the case in *Bank Overground* publications). Note that the object returned is a [gtable](https://gtable.r-lib.org/), rather than a ggplot, object, and can be drawn using standard functions from the `grid` package.

```{r title_align}
chart_two <- left_align_titles(chart_two)

grid::grid.draw(chart_two)
```


### Monetary Policy Report (MPR)

`theme_mpr` approximates the design of charts in [the Monetary Policy Report](https://www.bankofengland.co.uk/monetary-policy-report/2019/november-2019). Note the arguments passed to `scale_y_continuous()` to achieve the correct y-axis format (as well as `scale_colour_discrete_boe`, seen earlier, to do the colours).

```{r ir_theme}

chart +
  # apply boe palette
  scale_colour_discrete_boe(palette = "boe") +
  # add MPR theme
  theme_mpr() +
  # axis settings
  scale_y_continuous(
    position = "right", expand = c(0, 0), sec.axis = dup_axis(labels = NULL),
    breaks = boe_breaks_numeric(), limits = boe_limits_numeric()
    )
```

### DSD Publications

### Money & Credit

A theme that approximates the design of charts in [Money & Credit statistical releases](https://www.bankofengland.co.uk/statistics/money-and-credit/2018/may-2018) is found in `theme_mcg_pub()`.

```{r mcg_theme}

chart +
  # apply boe palette
  scale_colour_discrete_boe(palette = "boe") +
  # add MCG theme
  theme_mcg_pub() +
  # axis settings
  scale_y_continuous(position = "right", expand = c(0, 0), 
                     breaks = boe_breaks_numeric(), limits = boe_limits_numeric()) +
  scale_x_date(expand = c(0, 0), labels = boe_date_labels(),
               breaks = boe_breaks_date(), limits = boe_limits_date())
```
